Regularization Techniques for PSF--Matching Kernels. II. Pre--Filtering of Input Images

### 2015-02-12 submit/ms.tex
Normalizing PhoSim capitalization

We describe a novel technique for reducing the numbers of false detections in difference images due to deconvolution (or sharpening) by the profile--matching kernel.
To accomplish this, we pre--smooth the input science image by its own point spread function (PSF) before performing profile--matching.
This results in a maximum--likelihood image for point source detection, which has a $\sqrt{2}$ broader point--source profile than the image PSF.
We then match a reference image to this smoothed image, effectively resulting in a point--source maximum likelihood difference image.
%The process of detection and measurement happens directly on this difference image.
By smoothing the input image, we reduce the number of conditions under which the reference image has to be sharpened to match the input image, which is a significant source of systematic false detections.
We compared this pre--filtering technique with the traditional (post--filtering) image subtraction approach using a suite of high--fidelity simulated images from the Large Synoptic Survey Telescope (LSST) phoSim effort.
We used a configuration environment to tune each pipeline, using an objective function that minimizes the mean--square residuals in a control sample of sources in the images.
We then ran a suite of images through image subtraction and source detection and measurement, using a simulated deep template image and three simulated science images that differ primarily in seeing.
Situations that lead to a deconvolution of the template image are shown to have a lower sensitivity to true variability, and higher incidence of false detections, due to enhancement of the image noise.
In this regard, using the pre--filtering technique is clearly preferred.
Under smoothing of the template image, we find that the quality of the difference image pixels is very similar under the two approaches.
However, the propagated per--pixel variance varies significantly at the detection stage, leading to differences in the rate of false detections.
We find a single--parameter correction to the effective detection threshold that brings the run of false detections with detection threshold in--line with theoretical expectations.
We examined several sources for this discrepancy.
An over or under--estimate of the background is shown to bias the ratio of positive--going to negative--going false detections, but does not significantly effect the total numbers.
However, a misunderstanding of the image noise is shown to significantly impact the overall false detection rate.
We find that even a 2\% misestimate of the noise may change the false detection rate by a factor of 1.5--2 at the 5--sigma detection threshold.
Under both pipelines, the propagated variance in the resulting difference images is within 1\% of the empirical variance.
However, with when post--filtering the difference images for detection, the variance is underestimated by 4--5\% (when not deconvolving), leading to an increase in the rate of false detections.
This suggests that the multiple integral transforms being applied to images lead to significant misestimation of the variance, as the pixels become correlated and current algorithms do not track the pixel covariance during these operations.
Finally, we note that in order for there to be less than 1 statistical false detection per LSST difference image, the detection pipeline must operate at a threshold of 5.5--sigma under median (0.6'') seeing conditions.

>>>>

We describe a novel technique for reducing the numbers of false detections in difference images due to deconvolution (or sharpening) by the profile--matching kernel.
To accomplish this, we pre--smooth the input science image by its own point spread function (PSF) before performing profile--matching.
This results in a maximum--likelihood image for point source detection, which has a $\sqrt{2}$ broader point--source profile than the image PSF.
We then match a reference image to this smoothed image, effectively resulting in a point--source maximum likelihood difference image.
%The process of detection and measurement happens directly on this difference image.
By smoothing the input image, we reduce the number of conditions under which the reference image has to be sharpened to match the input image, which is a significant source of systematic false detections.
We compared this pre--filtering technique with the traditional (post--filtering) image subtraction approach using a suite of high--fidelity simulated images from the Large Synoptic Survey Telescope (LSST) PhoSim effort.
We used a configuration environment to tune each pipeline, using an objective function that minimizes the mean--square residuals in a control sample of sources in the images.
We then ran a suite of images through image subtraction and source detection and measurement, using a simulated deep template image and three simulated science images that differ primarily in seeing.
Situations that lead to a deconvolution of the template image are shown to have a lower sensitivity to true variability, and higher incidence of false detections, due to enhancement of the image noise.
In this regard, using the pre--filtering technique is clearly preferred.
Under smoothing of the template image, we find that the quality of the difference image pixels is very similar under the two approaches.
However, the propagated per--pixel variance varies significantly at the detection stage, leading to differences in the rate of false detections.
We find a single--parameter correction to the effective detection threshold that brings the run of false detections with detection threshold in--line with theoretical expectations.
We examined several sources for this discrepancy.
An over or under--estimate of the background is shown to bias the ratio of positive--going to negative--going false detections, but does not significantly effect the total numbers.
However, a misunderstanding of the image noise is shown to significantly impact the overall false detection rate.
We find that even a 2\% misestimate of the noise may change the false detection rate by a factor of 1.5--2 at the 5--sigma detection threshold.
Under both pipelines, the propagated variance in the resulting difference images is within 1\% of the empirical variance.
However, with when post--filtering the difference images for detection, the variance is underestimated by 4--5\% (when not deconvolving), leading to an increase in the rate of false detections.
This suggests that the multiple integral transforms being applied to images lead to significant misestimation of the variance, as the pixels become correlated and current algorithms do not track the pixel covariance during these operations.
Finally, we note that in order for there to be less than 1 statistical false detection per LSST difference image, the detection pipeline must operate at a threshold of 5.5--sigma under median (0.6'') seeing conditions.
